\documentclass[11pt,a4paper]{article}

\usepackage[margin=2.5cm]{geometry}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{graphicx}
\usepackage{float}
\usepackage{array}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{longtable}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    urlcolor=blue
}

\title{Porting Milvus GPU Vector Search to AMD\\[0.3em]
\large hipVS / hipRAFT Layer Plan (ROCm)}
\author{Konstantin Kolchin}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
This document describes a layered experiment plan for replacing NVIDIA cuVS/cuRAFT with
AMD hipVS/hipRAFT in the Milvus/Knowhere GPU path. It complements the CPU benchmarking
runbook (\texttt{ann\_framework\_runbook.tex}): Milvus Standalone CPU results on
\texttt{amd-rx7900xtx} are the control; GPU port work proceeds in five layers from ROCm sanity
through standalone libraries, Knowhere, Milvus source builds, and harness benchmarks.

\textbf{Status (2026-07-22):} Layers~\textbf{0--4 done} on the 7900 XTX lab host
(ROCm~7.0.2; hipVS IVF packer fixed; Knowhere HIP L2 search 108 assertions; DXC Knowhere
\texttt{2.5} merged; Milvus \texttt{v2.5.4} HIP Standalone sealed \texttt{GPU\_IVF\_FLAT}
and \texttt{GPU\_IVF\_PQ}; full SIFT-1M FLAT recall@10 $0.38\rightarrow0.98$; fair CPU vs GPU:
batched $\sim$5--7$\times$, VDBBench concurrent $\sim$2--4$\times$ at high \texttt{nprobe}).
Fair HIP vs CUDA on RTX~4080 (\texttt{milvusdb/milvus:v2.5.4-gpu}): FLAT and PQ
\texttt{m}=32 recall ladders \textbf{match} AMD. Library hipVS vs cuVS harness:
Section~\ref{sec:hipvs-vs-cuvs-lib}. See Sections~\ref{sec:layer-progress},
\ref{sec:layer4-pq}, \ref{sec:hip-vs-cuda}.
\end{abstract}

\tableofcontents
\clearpage

\section{Background}

Milvus does not call cuVS directly from Python. The stack is:

\begin{verbatim}
pymilvus / VectorDBBench
        |
   Milvus server
        |
   Knowhere  (vector engine: CPU + GPU indexes)
        |
   GPU path: GPU_IVF_FLAT, GPU_IVF_PQ, GPU_CAGRA, GPU_BRUTE_FORCE
        |
   NVIDIA: RAFT / cuVS
   AMD goal: hipRAFT + hipVS (ROCm-DS ports)
\end{verbatim}

\textbf{hipVS} is an AMD port of cuVS
(\url{https://github.com/ROCm-DS/hipVS},
\url{https://rocm.docs.amd.com/projects/hipVS/en/latest/}) with API compatibility as an
explicit goal. \textbf{hipRAFT} is the matching primitives layer (RAFT/cuRAFT analogue on
ROCm).

The stock \textbf{Milvus v2.5.4 CPU Docker image} used for CPU baselines does \emph{not}
exercise the GPU/HIP path. GPU experiments require \textbf{source builds} (or a custom GPU
image you build yourself). NVIDIA Milvus GPU images will not run on AMD.

\section{Prerequisites (CPU control)}

Before starting GPU layers, the following should be in place on the target AMD host
(see the main runbook for commands and measured tables):

\begin{itemize}[leftmargin=*]
  \item SIFT-1M HDF5 dataset (\texttt{data/sift-128-euclidean.hdf5}).
  \item Milvus Standalone Docker CPU (\texttt{milvus-docker/}, \texttt{user.yaml}
        \texttt{knowhere.simdType: sse4\_2}).
  \item Optional but recommended: VectorDBBench \texttt{milvusivfflat} IVF \texttt{nprobe}
        sweep as the same-host CPU Milvus control.
\end{itemize}

On \texttt{amd-rx7900xtx} (7900 XTX / \texttt{gfx1100}), CPU baselines and the IVF sweep
were captured 2026-06-30. GPU work does not replace those numbers; it adds a new backend
to compare against them.

\section{Layered strategy}

Do not attempt to ``replace cuVS in Milvus'' in one step. Use \textbf{five layers}
(Layer~0 through Layer~4), each with explicit pass/fail criteria. Only advance when the
current layer passes.

\subsection{Progress at a glance (lab host \texttt{amd-rx7900xtx})}
\label{sec:layer-progress}

\textbf{Completed (gate passed):} Layers~\textbf{0}, \textbf{1}, \textbf{1.5}, \textbf{2}, \textbf{3}, and \textbf{4}.

\begin{table}[h]
\centering
\caption{Layer status and pass criteria (as of 2026-07-17)}
\begin{tabular}{@{}cllp{0.38\textwidth}@{}}
\toprule
\textbf{Layer} & \textbf{Name} & \textbf{Status} & \textbf{Pass criterion / evidence} \\
\midrule
0 & ROCm + host sanity & \textbf{Done} & \texttt{rocminfo} gfx1100; Milvus CPU healthy \\
1 & hipRAFT / hipVS alone & \textbf{Done} & Build + \texttt{BRUTEFORCE\_C\_TEST}; IVF packer issues deferred to 1.5 \\
1.5 & IVF packer fix (gfx1100) & \textbf{Done} & Float IVF gtests 98/98 (2026-07-05) \\
2 & Knowhere + HIP & \textbf{Done} & Catch2 \texttt{Test Gpu Index Search}: 108 assertions (2026-07-09); DXC \texttt{2.5} merged \\
3 & Milvus source + GPU & \textbf{Done} & Sealed \texttt{GPU\_IVF\_FLAT}: IndexNode \texttt{GPU\_CUVS\_IVF\_FLAT}; QueryNode load OK (2026-07-17) \\
4 & Harness benchmarks & \textbf{Done} & Full SIFT-1M FLAT + PQ; HIP vs CUDA recall match (2026-07-22) \\
\bottomrule
\end{tabular}
\end{table}

\noindent\textit{Note:} Layer~2 ``Done'' means the Knowhere HIP L2-search gate passed. The full
Catch2 GPU suite still has known gaps (CAGRA bitset, IVF\_PQ TopK); those are not Layer~2
blockers. Layer~3 ``Done'' means Milvus HIP Standalone can \textbf{build and load} a sealed
\texttt{GPU\_IVF\_FLAT} index on gfx1100 (not a full-SIFT recall/QPS claim---that is Layer~4).

\section{Layer 0 --- Machine, ROCm, and CPU Milvus sanity}

\noindent\textit{Status: \textbf{Done} (see Section~\ref{sec:layer-progress}).}

\subsection{Goal}

Confirm that the AMD GPU stack and the existing CPU Milvus environment work \emph{before}
touching hipVS, Knowhere forks, or Milvus source builds.

\subsection{Target host (lab)}

\begin{tabularx}{\textwidth}{@{}>{\raggedright\arraybackslash}p{0.28\textwidth}X@{}}
\toprule
\textbf{Field} & \textbf{Value (7900 XTX lab)} \\
\midrule
Hostname & \texttt{amd-rx7900xtx} \\
CPU & AMD Ryzen 7 7700X \\
GPU & Radeon RX 7900 XTX (\texttt{gfx1100}) \\
iGPU & \texttt{gfx1036} (ignore for GPU benchmarks) \\
OS & Ubuntu 22.04 (Jammy) \\
ROCm & \texttt{7.0.2} (\url{https://repo.radeon.com/amdgpu-install/7.0.2/}); upgraded from \texttt{6.4.4} for Layer~1 \\
AMDGPU DKMS & \texttt{amdgpu/6.14.14} via \texttt{amdgpu-install --usecase=graphics,rocm} \\
Milvus & Standalone \texttt{v2.5.4} via \texttt{docker-compose} in \texttt{milvus-docker/} \\
\bottomrule
\end{tabularx}

\subsection{Install and verify ROCm}

Use the \textbf{official} AMDGPU + ROCm packages, not unrelated tools (e.g.\ \texttt{snap
install rocminfo}).

\begin{verbatim}
# Repos (example for ROCm 6.4.4 on Jammy)
# https://repo.radeon.com/amdgpu/6.4.4/ubuntu jammy
# https://repo.radeon.com/rocm/apt/6.4.4 jammy

sudo apt update
sudo apt install rocm amdgpu-dkms amdgpu-lib

# Reboot after first DKMS install if the driver was not loaded before
sudo reboot
\end{verbatim}

\textbf{Sanity commands} (record output in the teammate run template):

\begin{verbatim}
/opt/rocm/bin/rocminfo | grep -E 'Name:|Marketing|gfx'
/opt/rocm/bin/hipcc --version
/opt/rocm/bin/rocm-smi
groups | grep -E 'render|video'
uname -r
lspci | grep -i vga
lsmod | grep amdgpu
\end{verbatim}

\textbf{Pass:}
\begin{itemize}[leftmargin=*]
  \item \texttt{rocminfo} lists \texttt{Marketing Name: Radeon RX 7900 XTX} and agent
        \texttt{gfx1100}.
  \item No HIP runtime errors from \texttt{hipcc --version}.
  \item User is in \texttt{render} and/or \texttt{video} groups (or runs GPU jobs with
        appropriate device permissions).
\end{itemize}

\textbf{Measured on lab host (Layer~0 pass, later upgraded for Layer~1):}
\begin{verbatim}
rocminfo: gfx1100 / Radeon RX 7900 XTX visible
CPU agent: AMD Ryzen 7 7700X
Layer 0: ROCm 6.4.4; Layer 1 build: upgraded to ROCm 7.0.2 (see ROCm 7 upgrade subsection)
\end{verbatim}

Match hipVS/hipRAFT branch \texttt{release/rocmds-25.10} to \textbf{ROCm~$\geq$7.0.2} on this
host. ROCm~6.4.4 was insufficient for hipVS (missing hipCUB \texttt{DeviceFor}; see
Section~\ref{sec:layer1-rocm7-upgrade}).

\subsection{Upgrade to ROCm 7.0.2 (required for Layer~1 on gfx1100)}
\label{sec:layer1-rocm7-upgrade}

AMD does \textbf{not} support in-place ROCm upgrades. On Ubuntu~22.04 for Radeon + ROCm:

\begin{enumerate}[leftmargin=*]
  \item Install \texttt{dkms} and kernel headers if missing:
        \texttt{sudo apt install dkms linux-headers-\$(uname -r)}.
  \item Uninstall old stack: \texttt{sudo amdgpu-uninstall} (or \texttt{apt autoremove rocm*}).
  \item Remove stale repo files under \texttt{/etc/apt/sources.list.d/}; \texttt{apt clean}.
  \item Install \texttt{amdgpu-install} 7.0.2 \texttt{.deb} from AMD; run
        \texttt{sudo amdgpu-install -y --usecase=graphics,rocm}.
  \item Reboot; add user to \texttt{render,video}; verify \texttt{rocminfo}, \texttt{hipcc},
        and \texttt{/opt/rocm/include/hipcub/device/device\_for.hpp} exists.
\end{enumerate}

Full command transcript: AMD Radeon install guide for ROCm~7.0.2 on Jammy
(\url{https://rocm.docs.amd.com/projects/radeon-ryzen/en/docs-7.0.2/docs/install/installrad/native_linux/install-radeon.html}).

\subsection{Verify CPU Milvus (control stack)}

This host uses classic \texttt{docker-compose} (hyphen), not the Compose V2 plugin
\texttt{docker compose}.

\begin{verbatim}
cd ~/ann-harness-amd/milvus-docker
docker-compose up -d
sleep 15
docker-compose ps
curl -sf http://127.0.0.1:9091/healthz && echo OK
\end{verbatim}

\textbf{Pass:} \texttt{milvus-standalone} reports \texttt{Up (healthy)}; health endpoint
returns OK.

\subsection{Long runs over SSH}

Use \texttt{tmux} for any Layer~1+ build or benchmark. Detach with \textbf{Ctrl+B, D} before
closing SSH. An empty \texttt{tmux ls} after reboot is normal---start a new session
(\texttt{tmux new -s rocm-check}).

\subsection{Layer~0 checklist (host readiness)}

\begin{table}[h]
\centering
\caption{Host readiness checklist (\texttt{amd-rx7900xtx}); see also Section~\ref{sec:layer-progress}}
\begin{tabular}{@{}lr@{}}
\toprule
\textbf{Item} & \textbf{Status} \\
\midrule
ROCm installed (\texttt{7.0.2}) & Done \\
\texttt{rocminfo} shows \texttt{gfx1100} / 7900 XTX & Done \\
Milvus Standalone Docker healthy & Done \\
CPU Milvus VDBBench IVF baselines (runbook Tables~6--7) & Done \\
\bottomrule
\end{tabular}
\end{table}

\noindent Later layers (hipVS, packer, Knowhere, Milvus GPU) are tracked in
Section~\ref{sec:layer-progress}, not in this Layer~0-only checklist.

\section{Layer 1 --- hipRAFT and hipVS in isolation}
\label{sec:layer1-hipvs-hipraft}

\noindent\textit{Status: \textbf{Done} (build + brute-force; IVF packer completed in Layer~1.5).}

\subsection{Goal}

Prove AMD libraries work on the GPU \textbf{before} Knowhere or Milvus. Layer~1 success means:
\texttt{libcuvs.so} builds for \texttt{gfx1100}, hipVS tests run on GPU, and IVF/brute-force search
returns sensible neighbors.

\subsection{Reality check (7900 XTX / RDNA3)}

\begin{itemize}[leftmargin=*]
  \item \textbf{ROCm~$\geq$7.0.2 is required} for \texttt{release/rocmds-25.10} on this branch.
        ROCm~6.4.4 hipCUB lacks \texttt{cub::DeviceFor} (\texttt{device\_for.hpp}); hipVS
        configure/compile fails with \texttt{no member named 'DeviceFor' in namespace 'hipcub'}.
  \item \textbf{RDNA3 / \texttt{gfx1100} requires warp size 32.} hipCollections (cuco) defaults to
        CDNA-style warp size 64; without \texttt{USE\_WARPSIZE\_32=ON} configure fails with
        \texttt{not supported with warp size 64 (gfx1100)}. This path is \textbf{experimental}
        in upstream cuco.
  \item Official hipVS docs list \textbf{Instinct GPUs only} (\texttt{gfx90a}/\texttt{gfx942});
        \texttt{gfx1100} is community/experimental. Stock \texttt{release/rocmds-25.10} fails IVF
        packer gtests until \texttt{kIndexGroupSize=32} patch (Layer~1.5, Section~\ref{sec:layer15-measured}).
  \item First build downloads dependencies via ROCmDS-cmake and can take \textbf{1--3+ hours} and
        \textbf{10--30\,GB} disk under \texttt{\textasciitilde{}rocmds\_check\_gfx1100/}.
\end{itemize}

\subsection{Step 1 --- Host packages (once)}

On \texttt{amd-rx7900xtx}, inside \texttt{tmux}:

\begin{verbatim}
sudo apt update
sudo apt install -y git ninja-build ccache \
  build-essential pkg-config \
  libopenblas-dev libblas-dev liblapack-dev libsuitesparse-dev

# Ubuntu 22.04 ships cmake 3.22; hipVS wants >= 3.31
python3 -m pip install --user 'cmake>=3.31'
export PATH="$HOME/.local/bin:$PATH"
cmake --version   # must show >= 3.31

# ROCm tools on PATH
export PATH=/opt/rocm/bin:$PATH
export ROCM_PATH=/opt/rocm
hipcc --version
/opt/rocm/bin/rocminfo | grep -E 'gfx1100|7900'
\end{verbatim}

Optional: add user to GPU groups if builds/tests fail with permission errors:

\begin{verbatim}
sudo usermod -aG render,video "$USER"
# log out and back in
\end{verbatim}

\subsection{Step 2 --- Clone pinned sources}

Use branch \texttt{release/rocmds-25.10} for both repos (matched pair):

\begin{verbatim}
export WORKDIR=~/rocmds_check_gfx1100
export INSTALL_PREFIX=$WORKDIR/install
mkdir -p "$WORKDIR" "$INSTALL_PREFIX"

cd "$WORKDIR"
git clone -b release/rocmds-25.10 --depth 1 https://github.com/ROCm-DS/hipRaft.git hipRaft
git clone -b release/rocmds-25.10 --depth 1 https://github.com/ROCm-DS/hipVS.git hipVS
# If directories already exist, skip clone; use: git checkout release/rocmds-25.10
\end{verbatim}

\subsection{Step 2b --- Patch rocThrust version for ROCm 6.4.x only}

\textbf{Skip this step on ROCm~7.x} (system rocThrust is \textbf{4.0.0}; leave
\texttt{versions.json} unchanged).

On ROCm~6.4.4 only: hipRaft pins rocThrust \textbf{4.0.0} but apt ships \textbf{3.3.x}.
Without the patch, \texttt{find\_package} fails and CPM tries the sentinel git tag
\texttt{We-always-use-find\_package-for-rocthrust.} (which is not a real branch).

\begin{verbatim}
python3 <<'PY'
import json
from pathlib import Path
p = Path.home() / "rocmds_check_gfx1100/hipRaft/versions.json"
data = json.loads(p.read_text())
data.setdefault("packages", {}).setdefault("rocthrust", {})["version"] = "3.3.0"
p.write_text(json.dumps(data, indent=2) + "\n")
print("patched", p)
PY
\end{verbatim}

\subsection{Step 3 --- Environment for builds}

\begin{verbatim}
export PATH=$HOME/.local/bin:/opt/rocm/bin:$PATH
export ROCM_PATH=/opt/rocm-7.0.2
export WORKDIR=~/rocmds_check_gfx1100
export INSTALL_PREFIX=$WORKDIR/install
export PARALLEL_LEVEL=$(nproc)

export CMAKE_PREFIX_PATH=$ROCM_PATH:$ROCM_PATH/lib/cmake:$INSTALL_PREFIX/lib/cmake
export HIP_DIR=$ROCM_PATH/lib/cmake/hip
export rocthrust_DIR=$ROCM_PATH/lib/cmake/rocthrust
export rocprim_DIR=$ROCM_PATH/lib/cmake/rocprim
export hipcub_DIR=$ROCM_PATH/lib/cmake/hipcub
export RAPIDS_CMAKE_ROCTHRUST_USE_LOCAL=ON
export RAPIDS_CMAKE_ROCM_ORG=ROCm
export RAPIDS_CMAKE_ROCM_DS_ORG=ROCm-DS
\end{verbatim}

Do \textbf{not} pass \texttt{--cmake-args=-D...} (bash strips the inner quotes; \texttt{build.sh}
then fails \texttt{Invalid option}). For \texttt{gfx1100}, wrap the \textbf{entire} flag in single
quotes so literal \texttt{"..."} reach \texttt{build.sh} (Step~4).

\subsection{Step 4 --- Build hipRAFT first}

hipVS depends on hipRAFT. Install both into \texttt{\$INSTALL\_PREFIX}:

\begin{verbatim}
cd $WORKDIR/hipRaft
./build.sh clean && rm -rf cpp/build _deps
# --gpu-arch MUST be last. Single-quote the whole --cmake-args flag (bash quote trap).
INSTALL_PREFIX=$INSTALL_PREFIX ./build.sh libraft tests --compile-lib \
  --cache-tool=ccache \
  '--cmake-args="-DUSE_WARPSIZE_32=ON"' \
  --gpu-arch="gfx1100"
\end{verbatim}

\textbf{Pass:} exit code~0; libraries under \texttt{\$INSTALL\_PREFIX/lib} (e.g.\ \texttt{libraft.so}).

If configure complains about missing \texttt{hipblas}/\texttt{hipsparse}/\texttt{rocthrust}, install ROCm dev packages and export the \texttt{*\_DIR} variables from Step~3 (do not pass them via \texttt{--cmake-args}):

\begin{verbatim}
sudo apt install -y rocm-dev rocm-hip-sdk rocthrust-dev rocprim-dev hipcub-dev

# verify system rocThrust (must exist before hipRAFT configure)
ls /opt/rocm/lib/cmake/rocthrust/rocthrust-config.cmake \
   /opt/rocm-7.0.2/lib/cmake/rocthrust/rocthrust-config.cmake 2>/dev/null

export PATH=$HOME/.local/bin:/opt/rocm/bin:$PATH
export ROCM_PATH=/opt/rocm-7.0.2
export CMAKE_PREFIX_PATH=$ROCM_PATH:$ROCM_PATH/lib/cmake:$INSTALL_PREFIX/lib/cmake
export rocthrust_DIR=$ROCM_PATH/lib/cmake/rocthrust
export rocprim_DIR=$ROCM_PATH/lib/cmake/rocprim
export hipcub_DIR=$ROCM_PATH/lib/cmake/hipcub
export RAPIDS_CMAKE_ROCTHRUST_USE_LOCAL=ON
export RAPIDS_CMAKE_ROCM_ORG=ROCm

cd $WORKDIR/hipRaft
./build.sh clean && rm -rf cpp/build _deps
INSTALL_PREFIX=$INSTALL_PREFIX ./build.sh libraft tests --compile-lib \
  --cache-tool=ccache \
  '--cmake-args="-DUSE_WARPSIZE_32=ON"' \
  --gpu-arch="gfx1100"
\end{verbatim}

If you still see \texttt{Build step for rocthrust failed} or \texttt{github.com//rocThrust/},
patch \texttt{versions.json} (Step~2b), confirm \texttt{ls \$rocthrust\_DIR} succeeds, then wipe \texttt{cpp/build \_deps} and retry.

\subsection{Step 5 --- Build hipVS (libcuvs + focused tests)}

Start with a \textbf{small test subset} before compiling all gtests:

\begin{verbatim}
cd $WORKDIR/hipVS
./build.sh clean
INSTALL_PREFIX=$INSTALL_PREFIX CMAKE_PREFIX_PATH=$CMAKE_PREFIX_PATH \
  ./build.sh libcuvs tests --cache-tool=ccache \
  '--cmake-args="-DUSE_WARPSIZE_32=ON"' \
  --gpu-arch="gfx1100" \
  --limit-tests="NEIGHBORS_ANN_IVF_FLAT_TEST;BRUTEFORCE_C_TEST"
\end{verbatim}

Full test suite (much longer):

\begin{verbatim}
INSTALL_PREFIX=$INSTALL_PREFIX CMAKE_PREFIX_PATH=$CMAKE_PREFIX_PATH \
  ./build.sh libcuvs tests \
  '--cmake-args="-DUSE_WARPSIZE_32=ON"' \
  --gpu-arch="gfx1100"
cd $WORKDIR/hipVS/cpp/build
ctest --test-dir ./tests --output-on-failure
\end{verbatim}

\textbf{Pass:} \texttt{libcuvs.so} installed under \texttt{\$INSTALL\_PREFIX/lib}; at least
\texttt{BRUTEFORCE\_C\_TEST} exits~0 on \texttt{gfx1100}. Full
\texttt{NEIGHBORS\_ANN\_IVF\_FLAT\_TEST} may fail on RDNA3 despite good recall (see
Section~\ref{sec:ivf-gtest-decode}).

\subsection{Step 6 --- Run tests manually}

\begin{verbatim}
export HIP_VISIBLE_DEVICES=0
export LD_LIBRARY_PATH=$INSTALL_PREFIX/lib:$ROCM_PATH/lib:$LD_LIBRARY_PATH
GTEST=$WORKDIR/hipVS/cpp/build/gtests

# Layer 1 backstop (passed on lab host 2026-07-03)
$GTEST/BRUTEFORCE_C_TEST
echo "BRUTEFORCE exit: $?"

# IVF gtest — expect many FAIL on gfx1100; read log carefully (Section below)
$GTEST/NEIGHBORS_ANN_IVF_FLAT_TEST \
  --gtest_filter='AnnIVFFlatTest/AnnIVFFlatTestF_float.*' \
  2>&1 | tee $WORKDIR/ivf_flat_float_gtest.log

# Optional: float-only diagnostic; closer to Milvus GPU_IVF_FLAT (float vectors)
# Rebuild with --limit-tests="IVF_FLAT_C_TEST" first if not built
# $GTEST/IVF_FLAT_C_TEST

# watch GPU during tests (second terminal)
watch -n1 rocm-smi
\end{verbatim}

Logs from the repo helper script: \texttt{scripts/check\_rocmds\_gfx1100.sh} writes
\texttt{\$WORKDIR/logs/} and \texttt{SUMMARY.txt}.

\subsection{Step 7 --- One-shot script (recommended)}

From a fresh clone of \texttt{ann-harness-amd} on the lab host:

\begin{verbatim}
tmux new -s layer1
cd ~/ann-harness-amd
chmod +x scripts/check_rocmds_gfx1100.sh
# ensure cmake >= 3.31 and rocm on PATH (Step 1)
bash scripts/check_rocmds_gfx1100.sh 2>&1 | tee ~/rocmds_check_gfx1100/layer1_run.log
# Ctrl+B, D
\end{verbatim}

\subsection{Step 8 --- Optional Python API smoke test}

Only after C++ Layer~1 passes. Upstream recommends a \textbf{micromamba/conda} env; this is
optional for Milvus/Knowhere integration (C++ path is enough for Layer~2):

\begin{verbatim}
# See hipVS docs: conda/environments/all_rocm_arch-x86_64.yaml
# Build hipRAFT into env, then:
#   cd hipVS && ./build.sh libcuvs python
#   python -c "import cuvs; print('ok')"
\end{verbatim}

\subsection{Step 9 --- Recall sanity (float, small data)}

After Layer~1 builds pass, validate \textbf{float} IVF (what SIFT and \texttt{GPU\_IVF\_FLAT}
use---not \texttt{uint8} gtest dtypes) on a \textbf{10k-vector} slice from SIFT HDF5.
Use hipVS/gtest recall lines: values near $1.0$ mean neighbor search is sane; gross
divergence means wrong metric/arch. Full SIFT-1M GPU benchmarks belong in Layer~4.

Stock Milvus Docker cannot run \texttt{GPU\_IVF\_FLAT}; this step uses hipVS directly until
Layer~2--3. A harness script is not yet in-repo---use gtest recall lines as interim evidence,
then add a small hipVS or \texttt{cuvs} Python script when pursuing Knowhere integration.

\subsection{How to read \texttt{NEIGHBORS\_ANN\_IVF\_FLAT\_TEST} output}
\label{sec:ivf-gtest-decode}

Each gtest case (\texttt{AnnIVFFlat/NN}) runs three sub-tests in one \texttt{TEST\_P}:
\texttt{testIVFFlat()} (search/recall), \texttt{testPacker()} (GPU list packing math),
\texttt{testFilter()}. A single \texttt{[ FAILED ]} line does \textbf{not} mean IVF search is
broken.

\paragraph{Recall line (usually good news).}
\begin{verbatim}
[RAFT] [info] Recall = 1.000000 (10000/10000), ... below the threshold (eps = 0.001000).
\end{verbatim}
This checks \textbf{neighbor correctness}. Values near \textbf{1.0} mean GPU IVF search found
the right neighbors. This is what matters for SIFT / \texttt{GPU\_IVF\_FLAT}. \textbf{Do not}
try to fix failures by changing \texttt{eps}---recall often already passes.

\paragraph{Packer failure at \texttt{ann\_ivf\_flat.cuh:376} (fixable layout bug).}
\begin{verbatim}
Failure
Value of: thrust::reduce(..., mask, ...) == list_size * ps.dim
  Actual: false
  Expected: true
\end{verbatim}
This is \textbf{not} a recall threshold. It counts \texttt{true} bits in an internal packing
\texttt{mask} and expects \textbf{exactly} \texttt{list\_size $\times$ dim}. On \texttt{gfx1100}
with \texttt{USE\_WARPSIZE\_32=ON}, this assert fails for \textbf{both float and uint8} suites
even when recall is \textbf{1.0} (97/98 cases on the lab host, 2026-07-03).

\textbf{Interpretation:} search can pass while extend/pack layout invariants fail on some cluster
sizes. That is a real bug in the hipVS IVF interleaved layout path (or its test mask formula),
not a reason to ignore gtests. Upstream CI targets Instinct (\texttt{gfx90a}/\texttt{gfx942});
on RDNA3, \texttt{kIndexGroupSize = raft::warp\_size()} is \textbf{32}, not the fixed-32
assumption in much cuVS documentation tested mainly on warp-64 hardware. Fixing this in a
\textbf{project fork} is Layer~1.5 work (Section~\ref{sec:layer15-packer-fix}).

\paragraph{\texttt{GetParam()} decode.}
Example: \texttt{\{ 1000, 10000, 16, 10, 512, 1024, 0, 0,0,0 \}} means
\texttt{num\_queries=1000}, \texttt{num\_db\_vecs=10000}, \texttt{dim=16}, \texttt{k=10},
\texttt{nprobe=512}, \texttt{nlist=1024}, \texttt{metric=L2}, plus three booleans. The index
\texttt{/93} is just parameter set \#93, not an error code.

\paragraph{Do not weaken gtests to pass.} Lowering recall \texttt{eps} or commenting out the
packer \texttt{ASSERT\_TRUE} hides the layout bug. \textbf{Do} fix the library (or align the
mask formula with \texttt{codepacker::pack}) in a fork, then re-run gtests until
\texttt{NEIGHBORS\_ANN\_IVF\_FLAT\_TEST} is green. Application-level float recall on SIFT remains
the Milvus sign-off check in Layer~2--4.

\subsection{Layer~1 measured results (\texttt{amd-rx7900xtx}, 2026-07-03)}
\label{sec:layer1-measured}

\begin{table}[h]
\centering
\caption{Layer~1 outcomes on RX 7900 XTX after ROCm 7.0.2 upgrade}
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Item} & \textbf{Result} \\
\midrule
ROCm & 7.0.2; \texttt{hipcc} 7.0.51831; \texttt{device\_for.hpp} present \\
Branch & \texttt{release/rocmds-25.10} (hipRaft + hipVS) \\
Prefix & \texttt{\textasciitilde{}/rocmds\_check\_gfx1100/install} \\
Build flags & \texttt{USE\_WARPSIZE\_32=ON}, \texttt{--gpu-arch=gfx1100} \\
\texttt{libraft.so} & Built and installed \\
\texttt{libcuvs.so} & $\sim$254\,MB; links to ROCm~7 libs \\
\texttt{BRUTEFORCE\_C\_TEST} & \textbf{PASS} (2/2; recall 1.0) \\
\texttt{NEIGHBORS\_ANN\_IVF\_FLAT\_TEST} (float filter) & 97/98 \textbf{FAIL} on packer; recall often 1.0 \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Layer~1 baseline (2026-07-03):} hipRAFT + hipVS build; \texttt{BRUTEFORCE\_C\_TEST}
green. Stock hipVS float IVF gtest: 97/98 packer fail (recall often 1.0). \textbf{Layer~1.5
(2026-07-05):} patches \texttt{0001} (debug) + \texttt{0002} (\texttt{kIndexGroupSize=32});
98/98 pass---see Section~\ref{sec:layer15-measured}. Proceed to Layer~2.

\subsection{Troubleshooting}

\begin{table}[h]
\centering
\caption{Common Layer~1 failures}
\begin{tabular}{@{}p{0.32\textwidth}p{0.58\textwidth}@{}}
\toprule
\textbf{Symptom} & \textbf{What to try} \\
\midrule
cmake $<$ 3.31 & \texttt{pip install 'cmake>=3.31'}; verify \texttt{\textasciitilde{}/.local/bin/cmake} \\
Missing hipBLAS/hipSPARSE & \texttt{sudo apt install rocm-dev rocm-hip-sdk} \\
\texttt{no member named 'DeviceFor' in hipcub} & Upgrade to ROCm~$\geq$7.0.2; verify \texttt{device\_for.hpp} \\
\texttt{amdgpu dkms failed} / \texttt{dkms: command not found} & \texttt{apt install dkms linux-headers-\$(uname -r)}; re-run \texttt{amdgpu-install} \\
IVF gtest FAIL but Recall = 1.0 & Packer layout bug on gfx1100---apply patch \texttt{0002}; Section~\ref{sec:layer15-measured} \\
\texttt{github.com//rocThrust/} clone fail (ROCm 6.4) & Patch \texttt{versions.json} rocThrust to \texttt{3.3.0}; export \texttt{RAPIDS\_CMAKE\_ROCM\_ORG=ROCm} \\
\texttt{We-always-use-find\_package-for-rocthrust} & On ROCm~7: revert patch to \texttt{4.0.0}; install \texttt{rocthrust-dev}; set \texttt{rocthrust\_DIR}; wipe \texttt{\_deps} \\
\texttt{not supported with warp size 64 (gfx1100)} & Pass \texttt{'--cmake-args=\"-DUSE\_WARPSIZE\_32=ON\"'} before \texttt{--gpu-arch}; experimental RDNA3 path \\
\texttt{Invalid option: --cmake-args=...} & Bash ate inner quotes; use \texttt{'--cmake-args=\"-DFOO=bar\"'} not \texttt{--cmake-args=\"-DFOO=bar\"} \\
\texttt{Invalid option: --gpu-arch=...} & Put \texttt{--gpu-arch=\"gfx1100\"} \textbf{last}; do not omit it \\
HIP arch error / no kernel for GPU & Explicit \texttt{--gpu-arch="gfx1100"}; avoid \texttt{--allgpuarch} first \\
Segfault on RDNA3 & Reboot GPU (\texttt{sudo reboot}); retry; consider ROCm version pin per hipVS branch \\
OOM during build & Lower \texttt{PARALLEL\_LEVEL=8}; free disk; close Milvus if VRAM tight \\
\texttt{hnsw.cpp} / \texttt{AVX512Capable} & Stale Instinct cache or HNSW enabled; \texttt{./build.sh clean}; \texttt{rm -rf cpp/build}; add \texttt{-DBUILD\_CAGRA\_HNSWLIB=OFF}; \texttt{--gpu-arch=gfx1100} \\
\texttt{--offload-arch=gfx908} in compile line & Wipe \texttt{cpp/build}; reconfigure with \texttt{--gpu-arch=\"gfx1100\"} last \\
ROCm 6.4 vs docs 7.0.2 & Upgrade ROCm to match hipVS branch notes, or pass issue with build log \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Layer~1 pass criteria}

On \textbf{Instinct} (\texttt{gfx90a}/\texttt{gfx942}), upstream expects both gtests green. On
\textbf{RDNA3} (\texttt{gfx1100}):

\begin{itemize}[leftmargin=*]
  \item \textbf{Layer~1 (minimum):} hipRAFT \texttt{libraft.so} and hipVS \texttt{libcuvs.so} build
        with ROCm~$\geq$7.0.2; \texttt{BRUTEFORCE\_C\_TEST} exits~0.
  \item \textbf{Layer~1.5 (achieved 2026-07-05):} \texttt{NEIGHBORS\_ANN\_IVF\_FLAT\_TEST} float
        98/98 pass after \texttt{0002} patch (Section~\ref{sec:layer15-measured}).
  \item \texttt{rocm-smi} shows VRAM activity during tests.
  \item Record branch, fork commit, ROCm version, build flags, log paths (Table in
        Section~\ref{sec:layer1-measured}).
\end{itemize}

Repositories:
\begin{itemize}[leftmargin=*]
  \item \url{https://github.com/ROCm-DS/hipRaft} (branch \texttt{release/rocmds-25.10})
  \item \url{https://github.com/ROCm-DS/hipVS} (branch \texttt{release/rocmds-25.10})
  \item Build docs: \url{https://rocm.docs.amd.com/projects/hipVS/en/latest/install/build.html}
\end{itemize}

\textbf{Fail fast if:} hipVS lacks \texttt{GPU\_IVF\_FLAT}/brute-force coverage needed for your
first Knowhere target---check upstream before Layer~2 fork work.

\section{Layer 1.5 --- Fix IVF packer on \texttt{gfx1100} (fork work)}

\noindent\textit{Status: \textbf{Done} (2026-07-05; float IVF gtests 98/98).}
\label{sec:layer15-packer-fix}

\subsection{Goal}

Demonstrate that this project can \textbf{customize AMD ANN libraries}, not only consume them.
Layer~1.5 turns the observed \texttt{testPacker()} failures into a fork patch with green gtests
before Knowhere/Milvus integration.

\textbf{Pass:} \texttt{NEIGHBORS\_ANN\_IVF\_FLAT\_TEST} exits~0 on \texttt{gfx1100} (float suite
required; uint8 suite desirable). Record fork URL, commit hash, and log path in the runbook.

\subsection{Root-cause hypothesis (confirmed 2026-07-05)}

In hipVS \texttt{ivf\_flat.hpp}, interleaved cluster groups use
\texttt{kIndexGroupSize = raft::warp\_size()}. With \texttt{USE\_WARPSIZE\_32=ON} and
\texttt{--gpu-arch=gfx1100}, CUCO compiles for warp~32 but \texttt{raft::warp\_size()} still
resolved to \textbf{64} in headers---so \texttt{kIndexGroupSize=64} at runtime while packer
layout expected warp-32 grouping. Symptom: \texttt{mask\_true = 2 $\times$ expected} on case
\texttt{/0} (\texttt{list\_size=13}, \texttt{dim=1}); recall remained \textbf{1.0}.

\textbf{Fix (in-repo):} \texttt{patches/hipvs/0002-ivf-kIndexGroupSize-warp32-gfx1100.patch}
sets \texttt{kIndexGroupSize=32} when \texttt{CUCO\_USE\_WARPSIZE\_32} is defined. Apply via
\texttt{scripts/apply\_hipvs\_layer15\_patches.sh}.

\subsection{Step 1 --- Fork hipVS on GitHub and clone on the lab host}

\begin{enumerate}[leftmargin=*]
  \item On GitHub, fork \url{https://github.com/ROCm-DS/hipVS} to your account (e.g.\
        \url{https://github.com/konkolchin/hipVS}).
  \item On \texttt{amd-rx7900xtx}, clone the fork (or repoint an existing tree):
\end{enumerate}

\begin{verbatim}
export WORKDIR=~/rocmds_check_gfx1100
mkdir -p "$WORKDIR"
cd "$WORKDIR"

# fresh fork clone (recommended)
git clone -b release/rocmds-25.10 https://github.com/konkolchin/hipVS.git hipVS
cd hipVS
git remote add upstream https://github.com/ROCm-DS/hipVS.git

# create the Layer 1.5 working branch
git checkout -b gfx1100-ivf-packer-fix
\end{verbatim}

If you already built from \texttt{ROCm-DS/hipVS} under \texttt{hipVS/}, add your fork as
\texttt{origin} and push the branch there instead of re-cloning.

\subsection{Step 2 --- Apply Layer~1.5 patches from \texttt{ann-harness-amd}}

Patches live in this repo (no manual hipVS fork editing on the lab host unless you prefer it).
After \texttt{git pull} on \texttt{ann-harness-amd}, apply both patches to your hipVS tree:

\begin{verbatim}
cd ~/ann-harness-amd && git pull
bash scripts/apply_hipvs_layer15_patches.sh ~/rocmds_check_gfx1100/hipVS
# 0001: testPacker debug prints (mask_true vs expected)
# 0002: kIndexGroupSize=32 when CUCO_USE_WARPSIZE_32 (gfx1100 fix)
\end{verbatim}

Optional: commit on hipVS branch \texttt{gfx1100-ivf-packer-fix}:

\begin{verbatim}
cd ~/rocmds_check_gfx1100/hipVS
git checkout -b gfx1100-ivf-packer-fix 2>/dev/null || git checkout gfx1100-ivf-packer-fix
git add cpp/include/cuvs/neighbors/ivf_flat.hpp cpp/tests/neighbors/ann_ivf_flat.cuh
git commit -m "fix(ivf_flat): warp-32 kIndexGroupSize on gfx1100; packer debug"
\end{verbatim}

\subsection{Step 4 --- Rebuild, run gtest, and capture the first mismatch}
\label{sec:layer15-step4-run}

After applying the debug patch you \textbf{must rebuild}---edits to \texttt{ann\_ivf\_flat.cuh}
are not picked up until \texttt{NEIGHBORS\_ANN\_IVF\_FLAT\_TEST} is recompiled.

\textbf{Important:} wipe \texttt{cpp/build} before reconfigure. An old cache may still compile for
Instinct (\texttt{gfx908}/\texttt{gfx90a}/\texttt{gfx942}) and hit \texttt{hnsw.cpp} errors
(\texttt{AVX512Capable} with \texttt{-Werror}). IVF work does not need HNSW---disable it with
\texttt{-DBUILD\_CAGRA\_HNSWLIB=OFF}. Do \textbf{not} rely on \texttt{ninja NEIGHBORS\_ANN\_IVF\_FLAT\_TEST}
alone unless \texttt{CMakeCache.txt} already shows \texttt{gfx1100} only.

\begin{verbatim}
export PATH=$HOME/.local/bin:/opt/rocm/bin:$PATH
export ROCM_PATH=/opt/rocm-7.0.2
export WORKDIR=~/rocmds_check_gfx1100
export INSTALL_PREFIX=$WORKDIR/install
export CMAKE_PREFIX_PATH=$ROCM_PATH:$ROCM_PATH/lib/cmake:$INSTALL_PREFIX/lib/cmake
export rocthrust_DIR=$ROCM_PATH/lib/cmake/rocthrust
export rocprim_DIR=$ROCM_PATH/lib/cmake/rocprim
export hipcub_DIR=$ROCM_PATH/lib/cmake/hipcub
export RAPIDS_CMAKE_ROCTHRUST_USE_LOCAL=ON
export RAPIDS_CMAKE_ROCM_ORG=ROCm
export HIP_VISIBLE_DEVICES=0

# --- 4a. Clean + rebuild libcuvs + IVF gtest (required after patch) ---
# --compile-lib is hipRAFT-only; hipVS build.sh rejects it.
cd $WORKDIR/hipVS
./build.sh clean
rm -rf cpp/build

INSTALL_PREFIX=$INSTALL_PREFIX ./build.sh libcuvs tests \
  --cache-tool=ccache \
  '--cmake-args="-DUSE_WARPSIZE_32=ON -DBUILD_CAGRA_HNSWLIB=OFF"' \
  --gpu-arch="gfx1100" \
  --limit-tests="NEIGHBORS_ANN_IVF_FLAT_TEST"

grep CMAKE_HIP_ARCHITECTURES cpp/build/CMakeCache.txt
# expect: gfx1100 (not gfx908/gfx90a/gfx942)

ls -la $WORKDIR/hipVS/cpp/build/gtests/NEIGHBORS_ANN_IVF_FLAT_TEST
# Pass: build.sh exit 0 and binary exists

# --- 4b. Run float suite (matches SIFT / GPU_IVF_FLAT) ---
export LD_LIBRARY_PATH=$INSTALL_PREFIX/lib:$ROCM_PATH/lib:$LD_LIBRARY_PATH
GTEST=$WORKDIR/hipVS/cpp/build/gtests

$GTEST/NEIGHBORS_ANN_IVF_FLAT_TEST \
  --gtest_filter='AnnIVFFlatTest/AnnIVFFlatTestF_float.*' \
  2>&1 | tee $WORKDIR/packer_debug_float.log

echo "exit code: $?"
\end{verbatim}

\textbf{Pass for Step~4:} rebuild succeeds; log contains at least one
\texttt{testPacker mask mismatch} line (expected on stock \texttt{gfx1100} before the Step~5
fix). You do \textbf{not} need 98/98 green yet.

Extract the \textbf{first} debug line:

\begin{verbatim}
grep -m1 'testPacker mask mismatch' $WORKDIR/packer_debug_float.log
\end{verbatim}

Example line (yours will differ in \texttt{label} / sizes):

\begin{verbatim}
testPacker mask mismatch: label=17 list_size=142 dim=16 veclen=4
  kIndexGroupSize=32 padded_list_size=160 n_elems=2560
  mask_true=2288 expected=2272
\end{verbatim}

\paragraph{How to read the fields.}
\begin{table}[h]
\centering
\caption{Debug line field meanings (\texttt{testPacker})}
\begin{tabular}{@{}lp{0.62\textwidth}@{}}
\toprule
\textbf{Field} & \textbf{Meaning} \\
\midrule
\texttt{label} & IVF cluster (inverted list) index inside the built index \\
\texttt{list\_size} & Vectors in that cluster \\
\texttt{dim} & Vector dimension for this gtest case \\
\texttt{veclen} & Interleaved chunk size (\texttt{dim \% veclen == 0}) \\
\texttt{kIndexGroupSize} & Should be \textbf{32} on gfx1100 (\texttt{raft::warp\_size()}) \\
\texttt{padded\_list\_size} & \texttt{roundUp(list\_size)} to group boundary \\
\texttt{n\_elems} & \texttt{padded\_list\_size $\times$ dim} (buffer length) \\
\texttt{mask\_true} & Count of \texttt{true} bits in the test's validity mask \\
\texttt{expected} & \texttt{list\_size $\times$ dim} (must match for gtest pass) \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Direction of the bug.}
\begin{itemize}[leftmargin=*]
  \item \texttt{mask\_true > expected} --- mask marks too many buffer slots as valid (tail-group
        arithmetic in \texttt{ann\_ivf\_flat.cuh} is the prime suspect).
  \item \texttt{mask\_true < expected} --- mask too strict; compare against
        \texttt{pack\_1} indexing in \texttt{ivf\_flat\_helpers.cuh}.
  \item \texttt{kIndexGroupSize $\neq$ 32} --- rebuild flags wrong; fix Layer~1 env before
        editing source.
\end{itemize}

\paragraph{Map log line to a single gtest case.} Find the \texttt{[ FAILED ]} block nearest the
first mismatch; the suffix \texttt{/NN} is the parameter index:

\begin{verbatim}
grep -B5 -m1 'testPacker mask mismatch' $WORKDIR/packer_debug_float.log
# look for: [ FAILED ] AnnIVFFlatTest/AnnIVFFlatTestF_float.AnnIVFFlat/93
#            where GetParam() = { 1000, 10000, 16, 10, ... }
\end{verbatim}

Re-run \textbf{one} failing case for fast iteration (replace \texttt{/93} with your index):

\begin{verbatim}
$GTEST/NEIGHBORS_ANN_IVF_FLAT_TEST \
  --gtest_filter='AnnIVFFlatTest/AnnIVFFlatTestF_float.AnnIVFFlat/93' \
  2>&1 | tee $WORKDIR/packer_case_93.log
\end{verbatim}

Watch GPU activity in a second terminal: \texttt{watch -n1 rocm-smi}.

\subsection{Step 5 --- Fix in the fork, verify, push, record}
\label{sec:layer15-step5-fix}

Work on branch \texttt{gfx1100-ivf-packer-fix}. \textbf{Do not} disable the assert or lower recall
\texttt{eps}.

\paragraph{5a. Fix (in \texttt{ann-harness-amd} patches).} Step~4 showed \texttt{kIndexGroupSize=64}
despite \texttt{USE\_WARPSIZE\_32=ON}. Patch \texttt{0002-ivf-kIndexGroupSize-warp32-gfx1100.patch}
sets \texttt{kIndexGroupSize=32} when \texttt{CUCO\_USE\_WARPSIZE\_32} is defined (applied via
\texttt{scripts/apply\_hipvs\_layer15\_patches.sh}). If case \texttt{/0} still fails after
clean-rebuild, compare mask vs \texttt{pack\_1} as below.

\paragraph{5b. Compare test mask vs production packer (if still failing).} Open these side by side
for the failing \texttt{(list\_size, dim, veclen)} from Step~4:

\begin{itemize}[leftmargin=*]
  \item \texttt{cpp/tests/neighbors/ann\_ivf\_flat.cuh} --- \texttt{testPacker()} mask lambda
        (counts valid slots in the padded interleaved buffer).
  \item \texttt{cpp/src/neighbors/ivf\_flat/ivf\_flat\_helpers.cuh} ---
        \texttt{helpers::codepacker::pack\_1} / \texttt{unpack\_1} (production layout).
\end{itemize}

The test mask uses \texttt{surplus / chunk\_size} and \texttt{interleaved\_group::mod(...)};\
\texttt{pack\_1} uses \texttt{interleaved\_group::mod(offset) * veclen}. On tail groups
(\texttt{list\_size} not a multiple of \texttt{kIndexGroupSize}), those formulas must agree on
which of the \texttt{n\_elems} slots hold real data.

\paragraph{5c. Rebuild and verify loop.} After \texttt{apply\_hipvs\_layer15\_patches.sh}, clean-rebuild
and re-run case \texttt{/0}, then the full float filter:

\begin{verbatim}
cd $WORKDIR/hipVS
INSTALL_PREFIX=$INSTALL_PREFIX ./build.sh libcuvs tests \
  --cache-tool=ccache \
  '--cmake-args="-DUSE_WARPSIZE_32=ON -DBUILD_CAGRA_HNSWLIB=OFF"' \
  --gpu-arch="gfx1100" \
  --limit-tests="NEIGHBORS_ANN_IVF_FLAT_TEST"

export LD_LIBRARY_PATH=$INSTALL_PREFIX/lib:$ROCM_PATH/lib:$LD_LIBRARY_PATH
$GTEST/NEIGHBORS_ANN_IVF_FLAT_TEST \
  --gtest_filter='AnnIVFFlatTest/AnnIVFFlatTestF_float.AnnIVFFlat/93'
\end{verbatim}

\textbf{Case pass:} no \texttt{testPacker mask mismatch} line; no failure at
\texttt{ann\_ivf\_flat.cuh:376}; optional second assert (\texttt{devArrMatch} packed vs extend)
also passes.

\paragraph{5d. Full float suite, then uint8.}

\begin{verbatim}
$GTEST/NEIGHBORS_ANN_IVF_FLAT_TEST \
  --gtest_filter='AnnIVFFlatTest/AnnIVFFlatTestF_float.*' \
  2>&1 | tee $WORKDIR/packer_float_fixed.log
grep -c '\[  FAILED  \]' $WORKDIR/packer_float_fixed.log || true
# target: 0 failures, exit code 0

# optional (Milvus uint8 paths)
$GTEST/NEIGHBORS_ANN_IVF_FLAT_TEST \
  --gtest_filter='AnnIVFFlatTest/AnnIVFFlatTestF_uint8.*' \
  2>&1 | tee $WORKDIR/packer_uint8_fixed.log
\end{verbatim}

Confirm Layer~1 backstop still passes:

\begin{verbatim}
$GTEST/BRUTEFORCE_C_TEST
echo "BRUTEFORCE exit: $?"
\end{verbatim}

\paragraph{5e. Commit and push the fork.}

\begin{verbatim}
cd ~/rocmds_check_gfx1100/hipVS
git status
git add cpp/tests/neighbors/ann_ivf_flat.cuh \
        cpp/src/neighbors/ivf_flat/ivf_flat_helpers.cuh   # only files you changed
git commit -m "fix(ivf_flat): correct interleaved pack/mask on gfx1100 warp-32 tail groups"
git push -u origin gfx1100-ivf-packer-fix
git rev-parse HEAD   # copy this hash into the runbook
\end{verbatim}

Optional upstream PR: open from \texttt{konkolchin/hipVS:gfx1100-ivf-packer-fix} $\rightarrow$
\texttt{ROCm-DS/hipVS:release/rocmds-25.10} with the debug log, failing \texttt{GetParam()},
and the diff. Keep building from your fork until AMD merges.

\paragraph{5f. Record in \texttt{ann-harness-amd}.} Append to the lab log and update
Section~\ref{sec:layer1-measured} when done:

\begin{verbatim}
Layer 1.5 — amd-rx7900xtx — 2026-07-05
Patches: ann-harness-amd patches/hipvs/0001 + 0002 (apply_hipvs_layer15_patches.sh)
Fix: kIndexGroupSize=32 when CUCO_USE_WARPSIZE_32 (ivf_flat.hpp)
Before: kIndexGroupSize=64; case /0 mask_true=26 expected=13
NEIGHBORS_ANN_IVF_FLAT_TEST float: 98/98 PASS (~85s)
BRUTEFORCE_C_TEST: PASS
Logs: ~/rocmds_check_gfx1100/packer_float_fixed.log
       ~/rocmds_check_gfx1100/packer_case_0_fixed.log
\end{verbatim}

Reinstall into \texttt{\$INSTALL\_PREFIX} if you changed \texttt{libcuvs.so} after the last
\texttt{build.sh install} step, then pin this commit for Layer~2 Knowhere builds.

\subsection{Step 6 --- Layer~1.5 pass criteria}

\begin{itemize}[leftmargin=*]
  \item Fork pushed; commit hash recorded.
  \item \texttt{NEIGHBORS\_ANN\_IVF\_FLAT\_TEST} float: \textbf{98/98 PASS}, exit code~0.
  \item \texttt{BRUTEFORCE\_C\_TEST}: still PASS.
  \item \texttt{libcuvs.so} in \texttt{\$INSTALL\_PREFIX/lib} built from the fork commit.
\end{itemize}

Pin \texttt{INSTALL\_PREFIX/lib/libcuvs.so} to the patched build before Layer~2 Knowhere links against
hipVS.

\subsection{Layer~1.5 measured results (\texttt{amd-rx7900xtx}, 2026-07-05)}
\label{sec:layer15-measured}

\begin{table}[h]
\centering
\caption{Layer~1.5 outcomes after \texttt{kIndexGroupSize} patch on RX 7900 XTX}
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Item} & \textbf{Result} \\
\midrule
Patches & \texttt{ann-harness-amd} \texttt{patches/hipvs/0001} + \texttt{0002} \\
Apply script & \texttt{scripts/apply\_hipvs\_layer15\_patches.sh} \\
Fix & \texttt{kIndexGroupSize=32} when \texttt{CUCO\_USE\_WARPSIZE\_32} (\texttt{ivf\_flat.hpp}) \\
Build & \texttt{USE\_WARPSIZE\_32=ON}, \texttt{BUILD\_CAGRA\_HNSWLIB=OFF}, \texttt{--gpu-arch=gfx1100} \\
Before fix & \texttt{kIndexGroupSize=64}; case \texttt{/0}: \texttt{mask\_true=26}, \texttt{expected=13} \\
\texttt{NEIGHBORS\_ANN\_IVF\_FLAT\_TEST} (float) & \textbf{98/98 PASS}; exit 0; $\sim$85\,s \\
Case \texttt{/0} & \textbf{PASS}; recall 1.0 \\
\texttt{BRUTEFORCE\_C\_TEST} & \textbf{PASS} (unchanged) \\
Logs & \texttt{\textasciitilde{}/rocmds\_check\_gfx1100/packer\_float\_fixed.log} \\
\bottomrule
\end{tabular}
\end{table}

\section{Layer 2 --- Knowhere with HIP backend}
\label{sec:layer2-knowhere-hip}

\noindent\textit{Status: \textbf{Done} (2026-07-09; L2 search gate 108 assertions; DXC \texttt{2.5} merged).
Full GPU Catch2 suite still has known gaps---not Layer~2 blockers.}

\subsection{Goal}

Substitute the library link at the Knowhere layer, still without full Milvus: build
\texttt{libknowhere.so} with \texttt{WITH\_HIP=ON} against prebuilt hipVS/hipRAFT under
\texttt{\$INSTALL\_PREFIX}, and pass GPU unit tests for L2 search.

\subsection{Interaction diagram: Knowhere, hipVS, hipRAFT}

\begin{figure}[h]
\centering
\begin{tikzpicture}[
  node distance=1.2cm,
  box/.style={draw, rounded corners, align=center, minimum width=4.0cm, minimum height=0.95cm},
  smallbox/.style={draw, rounded corners, align=center, minimum width=3.7cm, minimum height=0.85cm},
  arr/.style={-{Latex[length=2.5mm]}, thick}
]
  \node[box] (milvus) {Milvus (QueryNode / IndexNode)\\create\_index(), search()};
  \node[box, below=of milvus] (knowhere) {Knowhere GPU wrappers\\GPU\_IVF\_FLAT / GPU\_BRUTE\_FORCE};
  \node[box, below=of knowhere] (hipvs) {hipVS ANN layer\\IVF build/search, top-k};
  \node[smallbox, below=of hipvs] (hipraft) {hipRAFT primitives\\distance, selection, kNN helpers};
  \node[smallbox, below=of hipraft] (rocm) {ROCm runtime + HIP kernels\\(gfx1100 device execution)};

  \draw[{Latex[length=2.5mm]}-{Latex[length=2.5mm]}, thick]
    (milvus) -- node[right, xshift=6pt, align=left] {\scriptsize index/search requests\\\scriptsize + result IDs/scores} (knowhere);
  \draw[{Latex[length=2.5mm]}-{Latex[length=2.5mm]}, thick]
    (knowhere) -- node[right, xshift=6pt, align=left] {\scriptsize API calls\\\scriptsize + results} (hipvs);
  \draw[{Latex[length=2.5mm]}-{Latex[length=2.5mm]}, thick]
    (hipvs) -- node[right, xshift=6pt, align=left] {\scriptsize primitive calls\\\scriptsize + outputs} (hipraft);
  \draw[{Latex[length=2.5mm]}-{Latex[length=2.5mm]}, thick]
    (hipraft) -- node[right, xshift=6pt, align=left] {\scriptsize kernel launches\\\scriptsize + distances/neighbors} (rocm);
\end{tikzpicture}
\caption{Layer~2 interaction model on AMD: Knowhere routes GPU index work to hipVS, hipVS uses hipRAFT primitives, and ROCm executes HIP kernels on the target GPU.}
\end{figure}

\subsection{Repos and branch}

\begin{tabularx}{\textwidth}{@{}>{\raggedright\arraybackslash}p{0.28\textwidth}X@{}}
\toprule
\textbf{Field} & \textbf{Value} \\
\midrule
DXC Knowhere & \url{https://github.dxc.com/llmkb-internal/knowhere} (branch \texttt{2.5}; PR~\#2 merged 2026-07-09) \\
Public mirror & \url{https://github.com/konkolchin/knowhere} branch \texttt{amd-hip-gfx1100-layer2} \\
Harness patches & \texttt{ann-harness-amd/patches/knowhere/0001}--\texttt{0047} \\
Build / test scripts & \texttt{scripts/build\_knowhere\_layer2.sh}, \texttt{scripts/run\_knowhere\_gpu\_tests.sh} \\
Lab tree & \texttt{\textasciitilde{}/rocmds\_check\_gfx1100/knowhere} \\
\bottomrule
\end{tabularx}

\subsection{Steps (reproducible)}

\begin{enumerate}[leftmargin=*]
  \item Ensure Layer~1.5 \texttt{libcuvs.so} / hipRAFT are installed under
        \texttt{\$INSTALL\_PREFIX} (patched packer).
  \item Clone Knowhere \texttt{2.5} (DXC tree preferred) or apply harness patches to
        upstream \texttt{zilliztech/knowhere} \texttt{2.5}:
\begin{verbatim}
bash scripts/apply_knowhere_layer2_patches.sh ~/rocmds_check_gfx1100/knowhere
\end{verbatim}
  \item Configure and build with Conan + HIP (not bare cmake into \texttt{build-hip}):
\begin{verbatim}
export WORKDIR=~/rocmds_check_gfx1100
export INSTALL_PREFIX=$WORKDIR/install
export ROCM_PATH=/opt/rocm
SKIP_GIT_PULL=1 bash scripts/build_knowhere_layer2.sh
\end{verbatim}
  \item Run the Layer~2 gate (Catch2 \textbf{section} under \texttt{Test All GPU Index};
        there is no section named ``\ldots{} L2 Metric''):
\begin{verbatim}
bash scripts/run_knowhere_gpu_tests.sh
# equivalent:
# knowhere_tests 'Test All GPU Index' --section 'Test Gpu Index Search'
\end{verbatim}
\end{enumerate}

\textbf{Pass (Layer~2):} section \texttt{Test Gpu Index Search} reports assertions $>0$ and
all green (L2 search over GPU/CUVS indexes including IVF\_FLAT / IVF\_PQ / CAGRA). Full
\texttt{Test All GPU Index} is \emph{not} the gate---see known gaps below.

This is where ``replace cuVS/cuRAFT'' actually happens---in Knowhere's build and GPU index
wrappers, not in \texttt{run\_milvus\_hdf5.py}.

\subsection{Measured on lab host (Layer~2 pass)}
\label{sec:layer2-measured}

\begin{table}[h]
\centering
\caption{Layer~2 results (\texttt{amd-rx7900xtx}, 2026-07-09)}
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Item} & \textbf{Result} \\
\midrule
ROCm / arch & 7.0.2 / \texttt{gfx1100} \\
Knowhere base & \texttt{2.5} + HIP CMake / shims (patches 0001--0047) \\
Build & \texttt{libknowhere.so} + \texttt{knowhere\_tests} via Conan + \texttt{WITH\_HIP} \\
Gate test & \texttt{Test Gpu Index Search} --- \textbf{108 assertions passed} \\
Runtime helpers & gflags shim (\texttt{libgflags\_gflagsns.so}); spdlog via \texttt{SPDLOG\_HEADER\_ONLY} \\
DXC merge & squash \texttt{bd82ff28} on \texttt{2.5} (PR~\#2) \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Known gaps on gfx1100 (not Layer~2 blockers)}

\begin{itemize}[leftmargin=*]
  \item \textbf{CAGRA + bitset:} recall $0.0$ in Bitset / Simple Bitset sections.
  \item \textbf{IVF\_PQ TopK:} occasional threshold miss (e.g.\ $0.8368 < 0.85$).
  \item Full suite may show $\sim$3--4 failed assertions among hundreds; treat as triage
        backlog, not a Layer~2 fail.
\end{itemize}

\section{Layer 3 --- Milvus from source (end-to-end)}
\label{sec:layer3-milvus-hip}

\noindent\textit{Status: \textbf{Done} (2026-07-17) --- sealed \texttt{GPU\_IVF\_FLAT} build/load on
\texttt{amd-rx7900xtx}; DXC branch \texttt{amd-hip-gfx1100-layer3} (Layer~4 green --- merge ready).}

\subsection{Goal}

Milvus Standalone uses the HIP-enabled Knowhere build (DXC \texttt{2.5} / local Layer~2 tree)
against prebuilt hipVS/hipRAFT. Stock Docker \texttt{milvusdb/milvus:v2.5.4} is CPU-only and
is \emph{not} the Layer~3 binary.

\subsection{Repos and scripts}

\begin{tabularx}{\textwidth}{@{}>{\raggedright\arraybackslash}p{0.28\textwidth}X@{}}
\toprule
\textbf{Field} & \textbf{Value} \\
\midrule
Milvus tag & \texttt{v2.5.4} (\texttt{milvus-io/milvus}) \\
Knowhere & Local \texttt{\$WORKDIR/knowhere} (HIP) or DXC \texttt{llmkb-internal/knowhere} \texttt{2.5} \\
hipVS/hipRAFT & \texttt{\$INSTALL\_PREFIX} (Layer~1.5) \\
DXC Milvus & \url{https://github.dxc.com/llmkb-internal/milvus} branch \texttt{amd-hip-gfx1100-layer3} \\
Harness patches & \texttt{patches/milvus/0001}--\texttt{0004} (Knowhere HIP FetchContent,
                  skip CUDA language, Knowhere API, clang~19 Tantivy) \\
Build & \texttt{scripts/build\_milvus\_layer3.sh} ($\rightarrow$ \texttt{make milvus-gpu}) \\
Smoke & \texttt{scripts/run\_milvus\_gpu\_smoke.sh} (passes \texttt{--flush}) \\
Client & \texttt{scripts/run\_milvus\_hdf5.py --index-type GPU\_IVF\_FLAT --flush} \\
PR notes & \texttt{docs/PR\_milvus\_amd\_hip\_gfx1100\_layer3.md} \\
\bottomrule
\end{tabularx}

\subsection{Steps (lab host, use \texttt{tmux})}

\begin{enumerate}[leftmargin=*]
  \item Confirm Layer~2 artifacts:
\begin{verbatim}
ls $HOME/rocmds_check_gfx1100/install/lib/libcuvs.so
ls $HOME/rocmds_check_gfx1100/knowhere/build/libknowhere.so
\end{verbatim}
  \item Pull harness and build Milvus (long):
\begin{verbatim}
cd ~/ann-harness-amd && git pull
export WORKDIR=~/rocmds_check_gfx1100
export INSTALL_PREFIX=$WORKDIR/install
export MILVUS_HIP_INSTALL_PREFIX=$INSTALL_PREFIX
export ROCM_PATH=/opt/rocm
# Pin discrete GPU (hide Raphael iGPU gfx1036)
export ROCR_VISIBLE_DEVICES=0 HIP_VISIBLE_DEVICES=0
# uses local HIP Knowhere when present; else FetchContent DXC 2.5
SKIP_CLONE=1 bash scripts/build_milvus_layer3.sh
\end{verbatim}
  \item Start HIP Milvus with writable rocksmq/runtime paths (not \texttt{/var/lib/milvus}),
        etcd/minio up, then smoke with \textbf{flush} so IndexNode seals GPU indexes:
\begin{verbatim}
bash scripts/run_milvus_gpu_smoke.sh
# or manually (unique collection avoids stale etcd node IDs):
# python scripts/run_milvus_hdf5.py --uri http://127.0.0.1:19530 \
#   --index-type GPU_IVF_FLAT --flush --nlist 128 --nprobes 8,16 \
#   --max-train-rows 50000 --max-query-rows 500 \
#   --collection sift_gpu_ivf_smoke_$(date +%s)
\end{verbatim}
\end{enumerate}

\textbf{Pass (Layer~3):}
\begin{itemize}[leftmargin=*]
  \item Client: insert $\rightarrow$ flush $\rightarrow$ \texttt{GPU\_IVF\_FLAT} index
        (\texttt{indexed\_rows} complete, \texttt{LoadState: Loaded}) $\rightarrow$ search.
  \item Logs: IndexNode writes \texttt{GPU\_CUVS\_IVF\_FLAT\_*}; QueryNode
        \texttt{CGO\_LOAD} \texttt{GPU\_CUVS\_IVF\_FLAT}; \textbf{no}
        \texttt{hipErrorInvalidDeviceFunction}.
  \item Do \emph{not} treat growing-path \texttt{IVF\_FLAT\_CC} remap alone as failure---that
        is the interim CPU index before seal. Without \texttt{--flush}, search often stays on
        \texttt{IVF\_FLAT\_CC} and is \emph{not} a HIP proof.
\end{itemize}

\subsection{Measured on lab host (Layer~3 pass)}
\label{sec:layer3-measured}

\begin{table}[h]
\centering
\caption{Layer~3 results (\texttt{amd-rx7900xtx}, 2026-07-17)}
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Item} & \textbf{Result} \\
\midrule
ROCm / arch & 7.0.2 / \texttt{gfx1100} (pin \texttt{ROCR\_VISIBLE\_DEVICES=0}) \\
Milvus & \texttt{v2.5.4} HIP Standalone (\texttt{bin/milvus}) \\
Knowhere & Nested Layer~2 HIP tree / DXC \texttt{2.5} \\
Index build & IndexNode: \texttt{GPU\_CUVS\_IVF\_FLAT\_0/1} + \texttt{SLICE\_META} \\
Sealed load & QueryNode \texttt{CGO\_LOAD} \texttt{GPU\_CUVS\_IVF\_FLAT}; no \texttt{InvalidDeviceFunction} \\
Client gate & 50k SIFT slice, \texttt{--flush}: \texttt{indexed\_rows=50000}, \texttt{Loaded} \\
DXC branch & \texttt{amd-hip-gfx1100-layer3} on \texttt{llmkb-internal/milvus} (open; merge after L4) \\
\bottomrule
\end{tabular}
\end{table}

\subsection{What the Milvus patches change}

\texttt{internal/core/thirdparty/knowhere/CMakeLists.txt} (\texttt{0001}):
\begin{itemize}[leftmargin=*]
  \item Default Knowhere remote $\rightarrow$ DXC \texttt{2.5} (override via CMake vars).
  \item Optional \texttt{MILVUS\_KNOWHERE\_SOURCE\_DIR} skips git and uses the Layer~2 tree.
  \item \texttt{MILVUS\_GPU\_VERSION=ON} forces \texttt{WITH\_CUVS} + \texttt{WITH\_HIP}
        (not NVIDIA \texttt{WITH\_RAFT}).
  \item Prefers \texttt{MILVUS\_HIP\_INSTALL\_PREFIX} / \texttt{ROCMDS\_INSTALL\_PREFIX}
        (not Milvus-clobbered \texttt{INSTALL\_PREFIX}) for hipVS/xxHash.
\end{itemize}
\texttt{internal/core/src/CMakeLists.txt} (\texttt{0002}):
\begin{itemize}[leftmargin=*]
  \item GPU build keeps \texttt{MILVUS\_GPU\_VERSION} but does \emph{not} enable the CUDA
        language (no \texttt{nvcc} on AMD).
\end{itemize}
\texttt{0003}/\texttt{0004}: Knowhere \texttt{IsAdditionalScalarSupported} API alignment;
clang~19 Tantivy \texttt{template} keyword fix.

\subsection{Known caveats (do not block Layer~3; watch in Layer~4)}

\begin{itemize}[leftmargin=*]
  \item \textbf{Dual GPU:} host has \texttt{gfx1100} + iGPU \texttt{gfx1036}. Without
        \texttt{ROCR\_VISIBLE\_DEVICES=0}, sealed deserialize can hit
        \texttt{hipErrorInvalidDeviceFunction}.
  \item \textbf{Growing segments} still remap \texttt{GPU\_IVF\_FLAT}$\rightarrow$\texttt{IVF\_FLAT\_CC};
        always \texttt{--flush} before claiming GPU search.
  \item \textbf{50k-slice recall} vs full-1M GT is meaningless; Layer~4 needs full train (or matching GT).
  \item \textbf{DXC merge:} keep \texttt{amd-hip-gfx1100-layer3} open until clean rebuild from the
        branch + full-SIFT evidence; optional Knowhere follow-up
        (\texttt{CMAKE\_CURRENT\_SOURCE\_DIR} for nested \texttt{.cu}).
\end{itemize}

\section{Layer 4 --- Harness benchmarks (comparison science)}
\label{sec:layer4-benchmarks}

\noindent\textit{Status: \textbf{Done} (2026-07-17) --- full SIFT-1M sealed \texttt{GPU\_IVF\_FLAT};
Table~\ref{tab:layer4-measured}. Checklist: \texttt{docs/layer4\_run\_checklist.md}.}

\subsection{Goal}

Measure sealed HIP \texttt{GPU\_IVF\_FLAT} against same-host CPU Milvus
(\texttt{IVF\_FLAT}) on SIFT-1M with the \textbf{same client protocol on both sides}.
Report two protocols separately (as in the management slides):

\begin{enumerate}[leftmargin=*]
  \item \textbf{Batched harness} --- one \texttt{search()} with all 10k queries
        (porting show-off / throughput).
  \item \textbf{VectorDBBench concurrent} --- industry tool, multi-client QPS.
\end{enumerate}

Never mix protocols in one speed-up column (that produced the withdrawn
$\sim$50$\times$ apples-to-oranges figures).

\begin{table}[h]
\centering
\caption{Layer~4 experiments}
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Experiment} & \textbf{Tool} & \textbf{What you learn} \\
\midrule
Batched CPU vs GPU & \texttt{run\_milvus\_hdf5.py} / Layer~4 & Same-protocol throughput speed-up \\
VDBBench concurrent & \texttt{milvusivfflat} / \texttt{milvusgpuivfflat} & Industry-tool QPS speed-up \\
HIP sealed smoke & \texttt{run\_milvus\_layer4.sh} & Engineering proof of \texttt{GPU\_CUVS} path \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Fair compare recipe (VectorDBBench)}

\begin{tabularx}{\textwidth}{@{}>{\raggedright\arraybackslash}p{0.28\textwidth}X@{}}
\toprule
\textbf{Field} & \textbf{Value} \\
\midrule
Dataset & SIFT-1M parquet (\texttt{\textasciitilde{}vdbbench-sift1m}) \\
Index & CPU \texttt{IVF\_FLAT} / GPU \texttt{GPU\_IVF\_FLAT}, \texttt{nlist=1024}, \texttt{k=10} \\
\texttt{nprobe} grid & $1,4,8,16,32$ \\
Runner & \texttt{scripts/run\_vdbbench\_milvus\_ivf\_sweep.sh} \\
Checklist & \texttt{docs/vdbbench\_cpu\_gpu\_compare.md} \\
Order & CPU Docker on \texttt{:19530}, then HIP Standalone on \texttt{:19530} \\
Outputs & \texttt{logs/vdb\_cpu\_ivf\_nprobe\_*.log}, \texttt{logs/vdb\_gpu\_ivf\_nprobe\_*.log} \\
\bottomrule
\end{tabularx}

\subsection{Steps (lab host, use \texttt{tmux})}

\begin{verbatim}
cd ~/ann-harness-amd && git pull
source ~/vdbbench-venv/bin/activate
# 1) Docker CPU Milvus on :19530
MODE=cpu bash scripts/run_vdbbench_milvus_ivf_sweep.sh
# 2) Stop Docker milvus; start HIP milvus (ROCR_VISIBLE_DEVICES=0)
MODE=gpu bash scripts/run_vdbbench_milvus_ivf_sweep.sh
\end{verbatim}

\textbf{Pass (Layer~4 fair compare):}
\begin{itemize}[leftmargin=*]
  \item Layer~4 batched GPU run finishes; recall@10 \textbf{rises} with \texttt{nprobe};
        fill Table~\ref{tab:layer4-cpu-gpu-batch} with same-protocol CPU batched QPS.
  \item VDBBench concurrent CPU + GPU sweeps finish; fill Table~\ref{tab:layer4-cpu-gpu}.
  \item GPU: sealed \texttt{GPU\_CUVS\_IVF\_FLAT} in milvus log; no \texttt{InvalidDeviceFunction}.
\end{itemize}

\subsection{Measured on lab host (Layer~4)}
\label{sec:layer4-measured}

Host: \texttt{amd-rx7900xtx}, Milvus \texttt{v2.5.4} HIP Standalone,
\texttt{ROCR\_VISIBLE\_DEVICES=0}, collection \texttt{sift\_gpu\_l4\_20260717\_200716},
\texttt{scripts/run\_milvus\_layer4.sh}.
Results JSON: \texttt{\$WORKDIR/logs/layer4\_gpu\_ivf\_20260717\_200716.json}.

Timings: insert $14.16$\,s, flush $3.02$\,s, index build $16.06$\,s;
\texttt{indexed\_rows=1000000}, load \texttt{Finished}/\texttt{Loaded}.
IndexNode saved multi-slice \texttt{GPU\_CUVS\_IVF\_FLAT\_*}; QueryNode
\texttt{CGO\_LOAD} \texttt{GPU\_CUVS\_IVF\_FLAT}; no \texttt{InvalidDeviceFunction}.

\begin{table}[h]
\centering
\caption{Layer~4 GPU Milvus SIFT-1M (\texttt{k=10}, \texttt{nlist=1024}, RX 7900 XTX, 2026-07-17)}
\label{tab:layer4-measured}
\begin{tabular}{@{}lrrr@{}}
\toprule
\textbf{nprobe} & \textbf{QPS} & \textbf{p99 ms} & \textbf{recall@10} \\
\midrule
1  & 16981.8 & 2.59 & 0.38 \\
4  & 21881.7 & 2.49 & 0.71 \\
8  & 21838.3 & 2.76 & 0.84 \\
16 & 26181.3 & 2.76 & 0.93 \\
32 & 19578.4 & 2.81 & 0.98 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Two fair CPU vs GPU comparisons (do not mix)}
\label{sec:layer4-cpu-gpu}

Same host \texttt{amd-rx7900xtx}, SIFT-1M L2, IVF \texttt{nlist=1024}, \texttt{k=10}.
Same order and numbers as the management slides.
Present \textbf{two} protocols separately; never divide batched GPU QPS by
serial VectorDBBench CPU QPS (that mix previously produced a spurious
$\sim$52$\times$ at \texttt{nprobe}=32 --- \textbf{withdrawn}).

\paragraph{Why VectorDBBench serial is not the speed headline.}
A serial VDBBench stage issues \textbf{one query per RPC} (latency/recall).
That under-utilizes GPU parallelism. Use it for recall only; see
\texttt{docs/vdbbench\_cpu\_gpu\_compare.md}. Throughput belongs in the
batched table below; industry multi-client QPS in the concurrent table.

\subsubsection{A --- Batched harness throughput (lead with this)}

Both sides: one \texttt{MilvusClient.search()} with all 10k queries
(\texttt{scripts/run\_milvus\_hdf5.py}). GPU = Table~\ref{tab:layer4-measured}.
CPU = same-host batched diagnostic (runbook Table~7);
re-run \texttt{scripts/run\_milvus\_batch\_cpu\_compare.sh} on Docker v2.5.4 for
strict same-day parity.

\begin{table}[H]
\centering
\caption{A: Milvus CPU vs GPU HIP --- batched 10k-query search (7900 XTX)
(same protocol both sides; matches slides ``Batched harness throughput'')}
\label{tab:layer4-cpu-gpu-batch}
\begin{tabular}{@{}lrrrr@{}}
\toprule
\textbf{nprobe} & \textbf{CPU QPS} & \textbf{GPU QPS} & \textbf{recall@10} & \textbf{QPS speed-up} \\
\midrule
1  & 22013 & 16982 & 0.38 & $0.77\times$ \\
4  & 14509 & 21882 & 0.71 & $1.51\times$ \\
8  &  9516 & 21838 & 0.84 & $2.30\times$ \\
16 &  5402 & 26181 & 0.93 & $4.85\times$ \\
32 &  2749 & 19578 & 0.98 & $7.12\times$ \\
\bottomrule
\end{tabular}
\end{table}

\noindent\textbf{Headline:} at \texttt{nprobe}=16--32, GPU HIP is $\sim$5--7$\times$ CPU
under the \textbf{same batched protocol}, with matching recall@10 (CPU and GPU agree
to two decimals) and sealed \texttt{GPU\_CUVS\_IVF\_FLAT}.

\subsubsection{B --- VectorDBBench concurrent}

\begin{table}[H]
\centering
\caption{B: Milvus CPU vs GPU HIP --- VectorDBBench concurrent QPS
(7900 XTX, 2026-07-18; concurrency $\{1,10,20,40,80\}$, 30s;
\texttt{vdb\_cpu\_concurrent\_nprobe\_20260718\_201956.log},
\texttt{vdb\_gpu\_concurrent\_nprobe\_20260718\_203847.log}).
recall@10 from serial companion (same index/\texttt{nprobe});
matches slides ``VectorDBBench concurrent''}
\label{tab:layer4-cpu-gpu}
\begin{tabular}{@{}lrrrr@{}}
\toprule
\textbf{nprobe} & \textbf{CPU QPS} & \textbf{GPU QPS} & \textbf{recall@10} & \textbf{QPS speed-up} \\
\midrule
1  & 13884 & 11110 & 0.38 & $0.80\times$ \\
4  & 10292 & 11066 & 0.71 & $1.08\times$ \\
8  &  7731 & 11110 & 0.84 & $1.44\times$ \\
16 &  4893 & 10904 & 0.93 & $2.23\times$ \\
32 &  2742 & 10374 & 0.98 & $3.78\times$ \\
\bottomrule
\end{tabular}
\end{table}

\noindent GPU HIP holds $\sim$11k QPS as \texttt{nprobe} rises; CPU falls off.
At high-recall operating points (\texttt{nprobe}=16--32): $\sim$2--4$\times$.
recall@10 (CPU and GPU) agree to two decimals.

\subsection{GPU\_IVF\_PQ on AMD (Layer~4 extension)}
\label{sec:layer4-pq}

Same host and batched harness as Table~\ref{tab:layer4-measured};
index \texttt{GPU\_IVF\_PQ}, \texttt{nlist}=1024, \texttt{nbits}=8, \texttt{k}=10
(2026-07-21). Sealed path: IndexNode \texttt{GPU\_CUVS\_IVF\_PQ}; QueryNode
\texttt{CGO\_LOAD} \texttt{GPU\_CUVS\_IVF\_PQ}.
Scripts: \texttt{scripts/run\_milvus\_layer4\_pq.sh} (\texttt{M=8|16|32}).
Plan notes: \texttt{docs/ivf\_pq\_amd\_cuda\_plan.md}.

Higher \texttt{m} $\Rightarrow$ less compression $\Rightarrow$ better recall
(still below FLAT --- expected). QPS stays in the same ballpark as FLAT
($\sim$15k--25k); the PQ product story is mainly memory / footprint.

\begin{table}[H]
\centering
\caption{AMD HIP \texttt{GPU\_IVF\_PQ} recall@10 vs FLAT
(SIFT-1M, \texttt{nlist}=1024, \texttt{nbits}=8, batched harness)}
\label{tab:layer4-pq-recall}
\begin{tabular}{@{}lrrrr@{}}
\toprule
\textbf{nprobe} & \textbf{m=8} & \textbf{m=16} & \textbf{m=32} & \textbf{FLAT} \\
\midrule
1  & 0.26 & 0.33 & 0.36 & 0.38 \\
4  & 0.35 & 0.50 & 0.60 & 0.71 \\
8  & 0.37 & 0.54 & 0.67 & 0.84 \\
16 & 0.38 & 0.56 & 0.71 & 0.93 \\
32 & 0.38 & 0.57 & 0.73 & 0.98 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{AMD HIP \texttt{GPU\_IVF\_PQ} QPS vs FLAT (same runs as Table~\ref{tab:layer4-pq-recall})}
\label{tab:layer4-pq-qps}
\begin{tabular}{@{}lrrrr@{}}
\toprule
\textbf{nprobe} & \textbf{m=8} & \textbf{m=16} & \textbf{m=32} & \textbf{FLAT} \\
\midrule
1  & 24882 & 16388 & 24273 & 16982 \\
4  & 21221 & 21084 & 20967 & 21882 \\
8  & 19766 & 19341 & 19500 & 21838 \\
16 & 23735 & 23566 & 22954 & 26181 \\
32 & 15402 & 15295 & 14784 & 19578 \\
\bottomrule
\end{tabular}
\end{table}

\noindent JSON: \texttt{layer4\_gpu\_ivf\_pq\_20260721\_213748.json} (\texttt{m}=8),
\texttt{\_213121.json} (\texttt{m}=16), \texttt{\_213909.json} (\texttt{m}=32)
under \texttt{\$WORKDIR/logs/}. Peer CUDA compare done for \texttt{m}=8, 16, and 32.

\subsection{Fair HIP vs CUDA (RTX 4080 peer)}
\label{sec:hip-vs-cuda}

Same client protocol as AMD Layer~4 (batched 10k-query \texttt{run\_milvus\_hdf5.py},
SIFT-1M, \texttt{nlist}=1024, \texttt{k}=10, \texttt{nprobe} grid, \texttt{--flush}).
CUDA host: RTX~4080, Ubuntu~22, \texttt{milvusdb/milvus:v2.5.4-gpu} via
\texttt{scripts/start\_milvus\_cuda\_gpu\_docker.sh}
(runbook: \texttt{docs/cuda\_4080\_benchmark\_runbook.md}).

\textbf{Hardware note:} RX~7900~XTX and RTX~4080 are peer-class enthusiast GPUs
(typical board power $\sim$320--355\,W). They are \emph{not} identical silicon;
use recall for port correctness and QPS for hardware/stack speed only.

\subsubsection{GPU\_IVF\_FLAT}

\begin{table}[H]
\centering
\caption{HIP vs CUDA --- \texttt{GPU\_IVF\_FLAT} recall@10 (batched; two decimals)}
\label{tab:hip-cuda-flat-recall}
\begin{tabular}{@{}lrr@{}}
\toprule
\textbf{nprobe} & \textbf{AMD HIP} & \textbf{CUDA 4080} \\
\midrule
1  & 0.38 & 0.38 \\
4  & 0.71 & 0.71 \\
8  & 0.84 & 0.84 \\
16 & 0.93 & 0.93 \\
32 & 0.98 & 0.98 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{HIP vs CUDA --- \texttt{GPU\_IVF\_FLAT} QPS
(speed-up $=$ AMD QPS / CUDA QPS)}
\label{tab:hip-cuda-flat-qps}
\begin{tabular}{@{}lrrr@{}}
\toprule
\textbf{nprobe} & \textbf{AMD HIP} & \textbf{CUDA 4080} & \textbf{Speed-up AMD vs NVIDIA} \\
\midrule
1  & 16982 & 11597 & $1.46\times$ \\
4  & 21882 & 19695 & $1.11\times$ \\
8  & 21838 & 19800 & $1.10\times$ \\
16 & 26181 & 18490 & $1.42\times$ \\
32 & 19578 & 13868 & $1.41\times$ \\
\bottomrule
\end{tabular}
\end{table}

\noindent CUDA JSON: \texttt{\textasciitilde{}milvus\_cuda\_4080/logs/layer4\_cuda\_gpu\_ivf\_flat\_20260722\_225252.json}.

\subsubsection{GPU\_IVF\_PQ (\texttt{m}=8)}

\begin{table}[H]
\centering
\caption{HIP vs CUDA --- \texttt{GPU\_IVF\_PQ} (\texttt{m}=8, \texttt{nbits}=8) recall@10}
\label{tab:hip-cuda-pq8-recall}
\begin{tabular}{@{}lrr@{}}
\toprule
\textbf{nprobe} & \textbf{AMD HIP} & \textbf{CUDA 4080} \\
\midrule
1  & 0.26 & 0.25 \\
4  & 0.35 & 0.35 \\
8  & 0.37 & 0.37 \\
16 & 0.38 & 0.38 \\
32 & 0.38 & 0.38 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{HIP vs CUDA --- \texttt{GPU\_IVF\_PQ} (\texttt{m}=8) QPS
(speed-up $=$ AMD QPS / CUDA QPS)}
\label{tab:hip-cuda-pq8-qps}
\begin{tabular}{@{}lrrr@{}}
\toprule
\textbf{nprobe} & \textbf{AMD HIP} & \textbf{CUDA 4080} & \textbf{Speed-up AMD vs NVIDIA} \\
\midrule
1  & 24882 & 15401 & $1.62\times$ \\
4  & 21221 & 19940 & $1.06\times$ \\
8  & 19766 & 20177 & $0.98\times$ \\
16 & 23735 & 19141 & $1.24\times$ \\
32 & 15402 & 14801 & $1.04\times$ \\
\bottomrule
\end{tabular}
\end{table}

\noindent CUDA JSON: \texttt{\textasciitilde{}milvus\_cuda\_4080/logs/layer4\_gpu\_ivf\_pq\_20260723\_190900.json}.

\subsubsection{GPU\_IVF\_PQ (\texttt{m}=16)}

\begin{table}[H]
\centering
\caption{HIP vs CUDA --- \texttt{GPU\_IVF\_PQ} (\texttt{m}=16, \texttt{nbits}=8) recall@10}
\label{tab:hip-cuda-pq16-recall}
\begin{tabular}{@{}lrr@{}}
\toprule
\textbf{nprobe} & \textbf{AMD HIP} & \textbf{CUDA 4080} \\
\midrule
1  & 0.33 & 0.33 \\
4  & 0.50 & 0.50 \\
8  & 0.54 & 0.54 \\
16 & 0.56 & 0.56 \\
32 & 0.57 & 0.57 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{HIP vs CUDA --- \texttt{GPU\_IVF\_PQ} (\texttt{m}=16) QPS
(speed-up $=$ AMD QPS / CUDA QPS)}
\label{tab:hip-cuda-pq16-qps}
\begin{tabular}{@{}lrrr@{}}
\toprule
\textbf{nprobe} & \textbf{AMD HIP} & \textbf{CUDA 4080} & \textbf{Speed-up AMD vs NVIDIA} \\
\midrule
1  & 16388 & 15540 & $1.05\times$ \\
4  & 21084 & 19716 & $1.07\times$ \\
8  & 19341 & 19829 & $0.98\times$ \\
16 & 23566 & 19155 & $1.23\times$ \\
32 & 15295 & 14717 & $1.04\times$ \\
\bottomrule
\end{tabular}
\end{table}

\noindent CUDA JSON: \texttt{\textasciitilde{}milvus\_cuda\_4080/logs/layer4\_gpu\_ivf\_pq\_20260723\_190933.json}.

\subsubsection{GPU\_IVF\_PQ (\texttt{m}=32)}

\begin{table}[H]
\centering
\caption{HIP vs CUDA --- \texttt{GPU\_IVF\_PQ} (\texttt{m}=32, \texttt{nbits}=8) recall@10}
\label{tab:hip-cuda-pq-recall}
\begin{tabular}{@{}lrr@{}}
\toprule
\textbf{nprobe} & \textbf{AMD HIP} & \textbf{CUDA 4080} \\
\midrule
1  & 0.36 & 0.36 \\
4  & 0.60 & 0.60 \\
8  & 0.67 & 0.67 \\
16 & 0.71 & 0.71 \\
32 & 0.73 & 0.73 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{HIP vs CUDA --- \texttt{GPU\_IVF\_PQ} (\texttt{m}=32) QPS
(speed-up $=$ AMD QPS / CUDA QPS)}
\label{tab:hip-cuda-pq-qps}
\begin{tabular}{@{}lrrr@{}}
\toprule
\textbf{nprobe} & \textbf{AMD HIP} & \textbf{CUDA 4080} & \textbf{Speed-up AMD vs NVIDIA} \\
\midrule
1  & 24273 & 11282 & $2.15\times$ \\
4  & 20967 & 19714 & $1.06\times$ \\
8  & 19500 & 19786 & $0.99\times$ \\
16 & 22954 & 18475 & $1.24\times$ \\
32 & 14784 & 14144 & $1.05\times$ \\
\bottomrule
\end{tabular}
\end{table}

\noindent CUDA JSON: \texttt{\textasciitilde{}milvus\_cuda\_4080/logs/layer4\_gpu\_ivf\_pq\_20260722\_225328.json}.

\paragraph{How to read HIP vs CUDA.}
Recall@10 matches on FLAT and PQ (\texttt{m}=8/16/32), so the HIP port returns the same
answer quality as CUDA on this recipe. QPS differences mainly reflect
\textbf{which GPU and software stack are faster}
(7900~XTX + HIP Standalone vs 4080 + CUDA Docker) --- not whether HIP search
returns wrong neighbors.

\subsection{Library-level hipVS vs cuVS (no Milvus)}
\label{sec:hipvs-vs-cuvs-lib}

Management ask: compare \textbf{hipVS} and \textbf{cuVS} speed as ANN libraries.
Section~\ref{sec:hip-vs-cuda} is the \emph{product} path (Milvus $\rightarrow$ Knowhere
$\rightarrow$ hipVS/cuVS). This subsection isolates the libraries.

\paragraph{Method.}
Shared Python API (\texttt{import cuvs}; hipVS keeps the cuVS package name).
Harness: \texttt{scripts/bench\_cuvs\_ivf.py} with wrappers
\texttt{run\_hipvs\_ivf\_bench.sh} / \texttt{run\_cuvs\_ivf\_bench.sh}.
Recipe locked to Layer~4: SIFT-1M HDF5, \texttt{nlist}=1024, \texttt{k}=10,
\texttt{nprobe}$\in\{1,4,8,16,32\}$, indexes \texttt{IVF\_FLAT} and
\texttt{IVF\_PQ} (\texttt{m}=32 $\mapsto$ cuVS \texttt{pq\_dim}, \texttt{nbits}=8
$\mapsto$ \texttt{pq\_bits}). Runbook: \texttt{docs/hipvs\_vs\_cuvs\_bench.md}.

\paragraph{Fairness.}
Same API and recipe; \textbf{different GPUs} (RX~7900~XTX vs RTX~4080).
Always report recall@10 with QPS. Speed-up $=$ hipVS QPS / cuVS QPS.
Compare JSON with \texttt{scripts/compare\_cuvs\_lib\_json.py}.

\paragraph{Status.}
Library \texttt{IVF\_PQ} (\texttt{m}=32) measured 2026-07-23 on both hosts
(JSON under \texttt{results/lib\_bench/} and \texttt{\$WORKDIR/logs/}).
Library \texttt{IVF\_FLAT} on AMD: batched search OK; single-query p99 path hits
\texttt{hipErrorInvalidValue} on gfx1100 --- re-run with \texttt{--p99-sample 0}.

\begin{table}[H]
\centering
\caption{Library hipVS vs cuVS --- \texttt{IVF\_PQ} (\texttt{m}=32, \texttt{nbits}=8)
recall@10 (batched; two decimals)}
\label{tab:lib-hipvs-cuvs-pq-recall}
\begin{tabular}{@{}lrr@{}}
\toprule
\textbf{nprobe} & \textbf{hipVS (7900 XTX)} & \textbf{cuVS (4080)} \\
\midrule
1  & 0.35 & 0.35 \\
4  & 0.59 & 0.59 \\
8  & 0.67 & 0.67 \\
16 & 0.71 & 0.71 \\
32 & 0.72 & 0.72 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{Library hipVS vs cuVS --- \texttt{IVF\_PQ} (\texttt{m}=32) QPS
(speed-up $=$ hipVS QPS / cuVS QPS)}
\label{tab:lib-hipvs-cuvs-pq-qps}
\begin{tabular}{@{}lrrr@{}}
\toprule
\textbf{nprobe} & \textbf{hipVS} & \textbf{cuVS} & \textbf{Speed-up} \\
\midrule
1  & 1593707 & 4326105 & $0.37\times$ \\
4  & 1268942 & 2297667 & $0.55\times$ \\
8  & 710989  & 1411722 & $0.50\times$ \\
16 & 393742  & 806267  & $0.49\times$ \\
32 & 217275  & 453316  & $0.48\times$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{How to read library vs product.}
Recall matches on the library path too. Raw library QPS is orders of magnitude
above Milvus (no RPC/Knowhere). On this PQ microbench, cuVS is about
$\sim$2$\times$ hipVS; under Milvus the same GPUs were $\sim$1$\times$
(Table~\ref{tab:hip-cuda-pq-qps}) because the service stack dominates.

\section{Metrics matrix}

For each backend (Milvus CPU IVF, Milvus GPU IVF), record:

\begin{table}[h]
\centering
\caption{Minimum metrics per experiment}
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Metric} & \textbf{Why} \\
\midrule
Index build time & GPU port often breaks here first \\
Load / cache time & \texttt{cache\_dataset\_on\_device} behavior \\
recall@10 vs \texttt{nprobe} & Correctness \\
QPS (batched 10k / VDBBench concurrent) & Same protocol on CPU and GPU; never mix \\
p99 latency & Serving angle \\
GPU memory peak & \texttt{rocm-smi} during build/search \\
Logs / stack traces & SIGILL, OOM, HIP errors \\
\bottomrule
\end{tabular}
\end{table}

Use one filled teammate run block per experiment (template in the main runbook appendix).

\section{Engineering workflow}

Recommended flow:

\begin{enumerate}[leftmargin=*]
  \item Layer~0: ROCm + Milvus CPU healthy on \texttt{amd-rx7900xtx}.
  \item Layer~1: hipVS/hipRAFT standalone on \texttt{gfx1100} (Section~\ref{sec:layer1-measured}).
  \item Layer~1.5: \texttt{kIndexGroupSize} patch; IVF gtests 98/98 (Section~\ref{sec:layer15-measured}).
  \item Layer~2: Knowhere + HIP; L2 GPU search UT green (Section~\ref{sec:layer2-measured}).
  \item Layer~3: Milvus source build; sealed \texttt{GPU\_IVF\_FLAT} build/load
        (Section~\ref{sec:layer3-measured}).
  \item Layer~4: same-protocol CPU vs GPU tables
        (batched Table~\ref{tab:layer4-cpu-gpu-batch}, then VDBBench concurrent
        Table~\ref{tab:layer4-cpu-gpu}).
  \item If recall OK and GPU faster at target points: widen index types + VectorDBBench GPU
        cases.
  \item If not: debug numerics, params, device cache before advancing.
\end{enumerate}

\textbf{Branching:} DXC Knowhere \texttt{2.5} (Layer~2 merged); Milvus branch
\texttt{amd-hip-gfx1100-layer3} on \texttt{llmkb-internal/milvus} --- ready to merge after
Layer~4 (Table~\ref{tab:layer4-measured}).
Pin versions: Milvus tag $\leftrightarrow$ Knowhere tag $\leftrightarrow$ hipVS
release $\leftrightarrow$ ROCm version in a single compatibility table per experiment.

\textbf{CI/local:} small dataset on every commit; full SIFT-1M nightly on the lab host in
\texttt{tmux}.

\section{Scope and realism}

\begin{itemize}[leftmargin=*]
  \item \textbf{hipVS is early} (e.g.\ v0.1.x)---expect gaps vs cuVS; start with
        \textbf{float} \texttt{GPU\_IVF\_FLAT} or GPU brute force, not every index at once.
  \item \textbf{RDNA3 IVF packer bug (fixed Layer~1.5)}---stock \texttt{release/rocmds-25.10} used
        \texttt{kIndexGroupSize=64} despite \texttt{USE\_WARPSIZE\_32}; patch \texttt{0002} in
        \texttt{ann-harness-amd} restores 98/98 float gtests on \texttt{gfx1100}.
  \item \textbf{API compatibility $\neq$ drop-in binary}---Milvus/Knowhere still need
        build-system and wrapper changes. Layers~2--3 deliver Knowhere HIP + Milvus
        Standalone sealed \texttt{GPU\_IVF\_FLAT}; Layer~4 measures end-to-end quality.
  \item \textbf{CPU baseline stays essential}---GPU work asks whether the hipVS path improves
        the recall--QPS trade-off vs CPU Milvus on the same host (same dataset and IVF recipe).
  \item \textbf{Stock Docker Milvus} remains the CPU control; GPU porting is source-build +
        fork work, not a config toggle.
\end{itemize}

\section{Immediate next steps (lab host)}

\textbf{Where we are:} Layers~0--4 complete (Section~\ref{sec:layer4-measured});
\texttt{GPU\_IVF\_PQ} measured on AMD (Section~\ref{sec:layer4-pq});
fair HIP vs CUDA on RTX~4080 done (Section~\ref{sec:hip-vs-cuda});
library hipVS vs cuVS harness ready (Section~\ref{sec:hipvs-vs-cuvs-lib}).
\textbf{What remains:} capture library JSON on both hosts; optional wider indexes;
public GitHub packaging.

\begin{enumerate}[leftmargin=*]
  \item \textbf{Library microbench:} on AMD and 4080 hosts run
        \texttt{bash scripts/run\_lib\_bench\_both\_indexes.sh}
        (runbook \texttt{docs/hipvs\_vs\_cuvs\_bench.md}); file JSON under
        \texttt{\$WORKDIR/logs/lib\_*.json}.
  \item Optional: \textbf{CAGRA} smoke + SIFT grid (watch Knowhere bitset gaps).
  \item Optional: VectorDBBench concurrent on CUDA for industry-tool HIP vs CUDA QPS.
  \item Copy / publish DXC HIP line notes to public GitHub when Business Value is agreed.
  \item Optional triage: CAGRA bitset recall$=0$; IVF\_PQ TopK threshold flakes on gfx1100
        (did not block sealed Layer~4 PQ on AMD).
\end{enumerate}

\section{References}

\begin{itemize}[leftmargin=*]
  \item Main CPU runbook: \texttt{docs/ann\_framework\_runbook.tex}
  \item Milvus Knowhere: \url{https://milvus.io/docs/knowhere.md}
  \item RAPIDS cuVS $\leftrightarrow$ Milvus: \url{https://docs.rapids.ai/api/cuvs/stable/integrations/milvus/}
  \item hipVS docs: \url{https://rocm.docs.amd.com/projects/hipVS/en/latest/}
  \item hipVS repo: \url{https://github.com/ROCm-DS/hipVS}
  \item hipRAFT repo: \url{https://github.com/ROCm-DS/hipRaft}
  \item DXC Knowhere (HIP): \url{https://github.dxc.com/llmkb-internal/knowhere}
  \item DXC Milvus (Layer~3 branch): \url{https://github.dxc.com/llmkb-internal/milvus}
  \item Public Knowhere mirror: \url{https://github.com/konkolchin/knowhere}
  \item Harness (patches + Layer~2/3/4 scripts): \url{https://github.com/konkolchin/ann-harness-amd}
  \item CUDA 4080 runbook: \texttt{docs/cuda\_4080\_benchmark\_runbook.md}
  \item hipVS vs cuVS library bench: \texttt{docs/hipvs\_vs\_cuvs\_bench.md}
  \item IVF\_PQ AMD$\leftrightarrow$CUDA plan: \texttt{docs/ivf\_pq\_amd\_cuda\_plan.md}
  \item Layer~2 PR notes: \texttt{docs/PR\_knowhere\_amd\_hip\_gfx1100.md}
  \item Layer~3 PR notes: \texttt{docs/PR\_milvus\_amd\_hip\_gfx1100\_layer3.md}
\end{itemize}

\end{document}
